### 2014-03-19 table_grammatical_paths.tex
I wanted to include this table, but it is probably hard to fit it in...

%
% File acl2014.tex
%
% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
%%
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2014}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{covington}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\begin{document}

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline \bf{Path} & \bf{Path group} & \bf{Score} \\ \hline
subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up & 3 & subj-pred \\
subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn & 3 & subj-pred \\
subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn{\textunderscore}{\textunderscore}INFV{\textunderscore}dn & 3 & subj-pred \\
subst:verb{\textunderscore}{\textunderscore}KOORD{\textunderscore}up{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up & 3 & subj-pred \\
subst:verb{\textunderscore}{\textunderscore}KOORD{\textunderscore}up{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn & 3 & subj-pred \\
subst:adj{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn & 3 & subj-pred \\
subst:adj{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn & 3 & subj-pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn & 3 & subj-pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn & 3 & subj-pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn & 3 & subj-within pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn & 3 & subj-within pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn{\textunderscore}{\textunderscore}KOORD{\textunderscore}dn & 3 & subj-within pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn{\textunderscore}{\textunderscore}DET{\textunderscore}dn & 3 & subj-within pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn{\textunderscore}{\textunderscore}DET{\textunderscore}dn & 3 & subj-within pred \\
subst:adj{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn{\textunderscore}{\textunderscore}ATR{\textunderscore}dn & 3 & subj-within pred \\
subst:adj{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn{\textunderscore}{\textunderscore}ATR{\textunderscore}dn & 3 & subj-within pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}ADV{\textunderscore}dn{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}dn & 2 & subj-within pred \\
subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn{\textunderscore}{\textunderscore}ADV{\textunderscore}dn{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}dn & 2 & subj-within pred \\
subst:adj{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}ADV{\textunderscore}dn & 2 & subj-within pred \\
subst:adj{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}INFV{\textunderscore}dn{\textunderscore}{\textunderscore}ADV{\textunderscore}dn & 2 & subj-within pred \\
subst:verb{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up & 1 & subj-within pred \\
subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}KOORD{\textunderscore}dn & 1 & subj-within pred \\
subst:verb{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up & 1 & pred-internal \\
subst:verb{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ADV{\textunderscore}up & 1 & pred-internal \\
subst:subst{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up{\textunderscore}{\textunderscore}ADV{\textunderscore}dn{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}dn & 1 & pred-internal \\
subst:adj{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up{\textunderscore}{\textunderscore}ADV{\textunderscore}dn & 1 & pred-internal \\
subst:adj{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ADV{\textunderscore}up & 1 & pred-internal \\
subst:verb{\textunderscore}{\textunderscore}DET{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up & 1 & pred-internal \\
subst:adj{\textunderscore}{\textunderscore}ATR{\textunderscore}dn & 2 & NP-internal \\
subst:subst{\textunderscore}{\textunderscore}ATR{\textunderscore}dn{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}dn & 2 & NP-internal \\
subst:subst{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up & 2 & NP-internal \\
subst:adj{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up & 2 & NP-internal \\
subst:verb{\textunderscore}{\textunderscore}ATR{\textunderscore}dn & 2 & NP-internal \\
subst:subst{\textunderscore}{\textunderscore}KOORD{\textunderscore}dn & 2 & NP-internal \\
subst:subst{\textunderscore}{\textunderscore}KOORD{\textunderscore}up & 2 & NP-internal \\
subst:subst{\textunderscore}{\textunderscore}DET{\textunderscore}dn & 2 & NP-internal \\
subst:subst{\textunderscore}{\textunderscore}DET{\textunderscore}up & 2 & NP-internal \\
subst:adj{\textunderscore}{\textunderscore}ATR{\textunderscore}dn{\textunderscore}{\textunderscore}ADV{\textunderscore}dn & 1 & NP-internal \\
subst:verb{\textunderscore}{\textunderscore}ATR{\textunderscore}dn{\textunderscore}{\textunderscore}KOORD{\textunderscore}dn & 1 & NP-internal \\
subst:adj{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}dn & 1 & NP-internal \\
subst:adj{\textunderscore}{\textunderscore}ATR{\textunderscore}dn{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}dn{\textunderscore}{\textunderscore}ATR{\textunderscore}dn & 1 & NP-internal \\
subst:subst{\textunderscore}{\textunderscore}KOORD{\textunderscore}up{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up & 3 & NP-internal \\
\hline
\end{tabular}
\end{center}
\caption{\label{grampath} Grammatical dependency paths }
\end{table}

\end{document}